\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2024-01-11 cubble.Rmd
modify affiliation and address for everyone

---
documentclass: jss
classoption: shortnames
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: 21 Chancellors Walk, Clayton VIC 3800 Australia
    email: dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of  Natural `\\`{=latex} Resources and Life Sciences \AND'
    address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU `\\`{=latex} United International College'
    address: 2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China
    email: nicolaslangrene@uic.edu.cn
  - name: Patricia Menéndez
    affiliation: University of Melbourne
    affiliation2: Monash University
    address: 813 Swanston Street, Parkville VIC 3052, Australia
    email:  Patricia.menendez@unimelb.edu.au
title:
  formatted: "\\pkg{cubble}: An \\proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
     Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, for instance, temporal trends of multiple variables across locations. In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on different spatio-temporal components.  The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis. In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling. The \code{cubble} class and the tools implemented in the package are illustrated with examples from climate data analysis.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial, temporal, spatio-temporal, "\\proglang{R}", environmental data, exploratory data analysis]
  plain:     [spatial, temporal, spatio-temporal, R, environmental data, exploratory data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
bibliography: references.bib
---

@@

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. The article is created using the package \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in \proglang{R} with the \code{rticles::jss_article} template. The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

>>>>

---
documentclass: jss
classoption: shortnames
author:
  - name: H. Sherry Zhang
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia
    email: \email{huize.zhang@monash.edu}
  - name: Dianne Cook
    affiliation: Monash University
    address: Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia
    email: dicook@monash.edu
  - name: Ursula Laa
    affiliation: 'University of  Natural `\\`{=latex} Resources and Life Sciences \AND'
    address: Institute of Statistics, University of  Natural Resources and Life Sciences, Peter-Jordan-Straße 82/I 1190 Vienna
    email:  ursula.laa@boku.ac.at
  - name: Nicolas Langrené
    affiliation: 'BNU-HKBU United International College'
    address:  Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College, Zhuhai 519087 China
    email: nicolaslangrene@uic.edu.cn
  - name: Patricia Menéndez
    affiliation: University of Melbourne
    affiliation2: Monash University
    address: School of Mathematics and Statistics, University of Melbourne, VIC 3052 Australia
    email:  Patricia.menendez@unimelb.edu.au
title:
  formatted: "\\pkg{cubble}: An \\proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
  # If you use tex in the formatted title, alsoge supply version without
  plain:     "cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
     Multivariate spatio-temporal data refers to multiple measurements taken across space and time. For many analyses, spatial and time components can be separately studied: for example, to explore the temporal trend of one variable for a single spatial location, or to model the spatial distribution of one variable at a given time. However for some studies, it is important to analyze different aspects of the spatio-temporal data simultaneously, for instance, temporal trends of multiple variables across locations. In order to facilitate the study of different portions or combinations of spatio-temporal data, we introduce a new class, \code{cubble}, with a suite of functions enabling easy slicing and dicing on different spatio-temporal components.  The proposed \code{cubble} class ensures that all the components of the data are easy to access and manipulate while providing flexibility for data analysis. In addition, the \pkg{cubble} package facilitates visual and numerical explorations of the data while easing data wrangling and modelling. The \code{cubble} class and the tools implemented in the package are illustrated with examples from climate data analysis.
keywords:
  # at least one keyword must be supplied
  formatted: [spatial, temporal, spatio-temporal, "\\proglang{R}", environmental data, exploratory data analysis]
  plain:     [spatial, temporal, spatio-temporal, R, environmental data, exploratory data analysis]
preamble: >
  \usepackage{amsmath}
  \usepackage{array}
output: 
  bookdown::pdf_book:
    base_format: rticles::jss_article
bibliography: references.bib
---

@@

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22. The article is created using the package \pkg{knitr} [@knitr] and \pkg{rmarkdown} [@rmarkdown] in \proglang{R} with the \code{rticles::jss_article} template. The source code for reproducing this paper can be found at: https://github.com/huizezhang-sherry/paper-cubble.

### 2024-01-11 cubble.tex
modify affiliation and address for everyone

\author{
H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU \\ United International College \And Patricia Menéndez\\University of Melbourne
}
\title{\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data}

@@

\Address{
    H. Sherry Zhang\\
    Monash University\\
    21 Chancellors Walk, Clayton VIC 3800 Australia\\
  E-mail: \email{huize.zhang@monash.edu}\\

      Dianne Cook\\
    Monash University\\
    21 Chancellors Walk, Clayton VIC 3800 Australia\\
  E-mail: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}\\

      Ursula Laa\\
    University of Natural \\ Resources and Life Sciences\\
    Gregor-Mendel-Straße 33, 1180 Wien, Austria\\
  E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\

      Nicolas Langrené\\
    BNU-HKBU \\ United International College\\
    2000 Jintong Road, Tangjiawan, Zhuhai, Guangdong Province, China\\
  E-mail: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}\\

      Patricia Menéndez\\
    Monash University\\
    813 Swanston Street, Parkville VIC 3052, Australia\\
  E-mail: \href{mailto:Patricia.menendez@unimelb.edu.au}{\nolinkurl{Patricia.menendez@unimelb.edu.au}}\\

@@

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. The article is created using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article} template. The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.

>>>>

\author{
H. Sherry Zhang\\Monash University \And Dianne Cook\\Monash University \And Ursula Laa\\University of Natural \\ Resources and Life Sciences \AND Nicolas Langrené\\BNU-HKBU United International College \And Patricia Menéndez\\University of Melbourne
}
\title{\pkg{cubble}: An \proglang{R} Package for Organizing and Wrangling Multivariate Spatio-temporal Data}

@@

\Address{
    H. Sherry Zhang\\
    Monash University\\
    Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia\\
  E-mail: \email{huize.zhang@monash.edu}\\

      Dianne Cook\\
    Monash University\\
    Department of Econometrics and Business Statistics, Monash University, VIC 3800 Australia\\
  E-mail: \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}}\\

      Ursula Laa\\
    University of Natural \\ Resources and Life Sciences\\
    Institute of Statistics, University of Natural Resources and Life Sciences, Peter-Jordan-Straße 82/I 1190 Vienna\\
  E-mail: \href{mailto:ursula.laa@boku.ac.at}{\nolinkurl{ursula.laa@boku.ac.at}}\\

      Nicolas Langrené\\
    BNU-HKBU United International College\\
    Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science, BNU-HKBU United International College, Zhuhai 519087 China\\
  E-mail: \href{mailto:nicolaslangrene@uic.edu.cn}{\nolinkurl{nicolaslangrene@uic.edu.cn}}\\

      Patricia Menéndez\\
    Monash University\\
    School of Mathematics and Statistics, University of Melbourne, VIC 3052 Australia\\
  E-mail: \href{mailto:Patricia.menendez@unimelb.edu.au}{\nolinkurl{Patricia.menendez@unimelb.edu.au}}\\

@@

This work is funded by a Commonwealth Scientific and Industrial Research Organisation (CSIRO) Data61 Scholarship and started while Nicolas Langrené was affiliated with CSIRO's Data61. Nicolas Langrené acknowledges the partial support of the Guangdong Provincial Key Laboratory IRADS (2022B1212010006, R0400001-22) and the UIC Start-up Research Fund UICR0700041-22. The article is created using the package \pkg{knitr} \citep{knitr} and \pkg{rmarkdown} \citep{rmarkdown} in \proglang{R} with the \code{rticles::jss_article} template. The source code for reproducing this paper can be found at: \url{https://github.com/huizezhang-sherry/paper-cubble}.
